DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Had to refactor all the functions not to use table to handle the degenerate case where the predictions were all 0s or 1s, so the table was not 2x2 but 2x1. Noe ROC should work

# 1. The Data
```{r loadData, echo=FALSE}
library(tidyverse)
# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

@@

```{r dataCM, echo=FALSE}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
```

@@

# 3. Accuracy
The function `ACC` below takes as input a data frame whose first two columns are
the true class and the modeled class respectively. It returns the accuracy of
the predictions using the following formula:

@@

```{r accuracy}
ACC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

@@

# 4. Classification Error Rate
The function `CER` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
classification error rate of the predictions using the following formula:

@@

```{r error}
CER <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    (sum(dt) - sum(diag(dt))) / sum(dt)
  }
}
```

@@

# 5. Precision
The function `PRC` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
classification error rate of the predictions using the following formula:

@@

```{r precision}
PRC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[2, 2] / sum(dt[, 2])
  }
}
```

@@

# 6. Sensitivity or Recall
The function `SNS` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
sensitivity of the predictions using the following formula:

@@

```{r sensitivity}
SNS <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[2, 2] / sum(dt[2, ])
  }
}
```

@@

# 7. Specificity
The function `SPC` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
specificity of the predictions using the following formula:

@@

```{r specificity}
SPC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[1, 1] / sum(dt[1, ])
  }
}
```

@@

# 8. \(F_1\) Score
The function `F1` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
\(F_1\) score of the predictions using the following formula:

@@

>>>>

# 1. The Data
```{r loadData, echo=FALSE}
# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

@@

```{r dataCM, echo=FALSE}
class_data <-  data.frame(class = data$class,
                          pred_class = data$scored.class,
                          pred_prob = data$scored.probability)
knitr::kable(table(class_data[,1:2]), caption = "Raw Confusion Matrix")
```

@@

# 3. Accuracy
The function `ACC` below takes as input a data frame whose first two columns are
the true class and the modeled class and named `class` and `pred_class`
respectively. It returns the accuracy of the predictions using the following
formula:

@@

```{r accuracy}
ACC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    sum(df$class == df$pred_class) / dim(df)[[1]]
  }
}
```

@@

# 4. Classification Error Rate
The function `CER` below takes as input a data frame whose first two columns
are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
classification error rate of the predictions using the following formula:

@@

```{r error}
CER <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    sum(df$class != df$pred_class) / dim(df)[[1]]
  }
}
```

@@

# 5. Precision
The function `PRC` below takes as input a data frame whose first two columns
are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
classification error rate of the predictions using the following formula:

@@

```{r precision}
PRC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    sum(df$class == 1 & df$pred_class == 1) / sum(df$pred_class == 1)
  }
}
```

@@

# 6. Sensitivity or Recall
The function `SNS` below takes as input a data frame whose first two columns
are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
sensitivity of the predictions using the following formula:

@@

```{r sensitivity}
SNS <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    sum(df$class == 1 & df$pred_class == 1) / sum(df$class == 1)
  }
}
```

@@

# 7. Specificity
The function `SPC` below takes as input a data frame whose first two columns
are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
specificity of the predictions using the following formula:

@@

```{r specificity}
SPC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    sum(df$class == 0 & df$pred_class == 0) / sum(df$class == 0)
  }
}
```

@@

# 8. \(F_1\) Score
The function `F1` below takes as input a data frame whose first two columns
are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
\(F_1\) score of the predictions using the following formula:

@@

10. ROC Curve
The ROC curve is defined as a plot of the true positive rate against the false
positive rate at different probability thresholds. The false positive rate is
usually plotted on the x-axis.

```{r ROC}
ROC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    rocTable <- data.table(thresh = seq(0, 1, 0.01),
                           x = double(101),
                           y = double(101))
    for (i in seq_along(rocTable$thresh)) {
      dftmp <- data.frame(class = df$class,
                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
                                              1, 0))
      rocTable$x[i] <- 1 - SPC(dftmp)
      rocTable$y[i] <- SNS(dftmp)
    }
    plot(rocTable$x, rocTable$y, type = 'l')
  }
}
```
